Formal Verification Tool Based on Symbolic Execution for Smart Contract

### 2023-11-13 papers/static-checker/ecoop2024/ecoop2024.tex
update the execution model

\usepackage{microtype}%if unwanted, comment out or use option "draft"

@@

\section{Executive Summary}
\summaryauthor[T.T.Ha Doan and P. Thiemann]{%
T.T.Ha Doan (University of Freiburg, Germany, doanha@informatik.uni-freiburg.de)\\
P. Thiemann (University of Freiburg, Germany, \\thiemann@informatik.uni-freiburg.de)
}

@@

This summary summarizes the outcomes of  our seminar. The seminar focused on\begin{itemize}
\item important issues,
\item relevant problems, and
\item adequate solutions.
\end{itemize}

As a major result from the seminar, the following problems have been identified: 
\begin{enumerate}
\item The problem of writing a brief, but concise executive summary.
\item The problem of collecting all abstracts from talks.
\item The problem of preparing summaries from working groups, open problem sessions, and panel discussions.
\end{enumerate}

@@

\section{Introduction}

\end{document}

>>>>

\usepackage{microtype}%if unwanted, comment out or use option "draft"
\pdfoutput=1
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathpartir}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{comment}
\usepackage{todonotes}
\newcommand\pt{\todo[author=PT,inline]}
\lstdefinelanguage{michelson}{
  basicstyle=\fontsize{8}{9.6}\selectfont,
  morekeywords={parameter,storage,or,unit,mutez,pair,bool,address}, sensitive=false,
  morecomment=[l]{\#},
  morecomment=[\STACK]{/*}{*/},
  morestring=[b]",
}
\lstset{
  language=Caml,
  captionpos=b,
  aboveskip=-\smallskipamount,
  belowskip=-\smallskipamount,
  belowcaptionskip=0pt,
  basicstyle=\fontsize{8}{9.6}\selectfont,
  morekeywords={val}
}

\input{macros}

@@

\summaryauthor[T.T.Ha Doan and P. Thiemann]{%
T.T.Ha Doan (University of Freiburg, Germany, doanha@informatik.uni-freiburg.de)\\
P. Thiemann (University of Freiburg, Germany, \\thiemann@informatik.uni-freiburg.de)
}

@@

@@

\section{Introduction}
\label{sec:introduction}

\section{Symblic Execution Model}
\label{sec:symblic-execution-model}

Let $\STACK = (\TermOne, \TYF) \STACKCONCAT (\TermTwo, \TYS) \STACKCONCAT \DOT
\STACKCONCAT \EMPTYSTACK$ be a stack, where elements are types paired with terms.
\\
Let $\INSTRUCTION = \InstructionOne; \InstructionTwo; \DOT; \InstructionN$ be a sequence of instructions. 
\\
Let \PREDICATE\ be a predicate.

\begin{definition}
A system state of the symbolic execution is a tuple \\ $\STATE = [\INSTRUCTION, \STACK, \PREDICATE]$. Let $\SYSTEM = \{\STATEONE, \STATETWO, \DOT, \STATEN \}$ range over sets of system states.
\end{definition}

Let $S_{init}$   = (\KPAIR\ \VPAR\ \VSTORAGE, \TPAIR\ \TYF\ \TYS) \STACKCONCAT\EMPTYSTACK\ be the initial  stack, where \VPAR\ and \VSTORAGE\ are the terms that represent the parameter and the storage and \TYF\ and \TYS\ are their corresponding types, and  $S_{final}$   = (\PAIR\ \VOPERATIONLIST\ \VSTORAGE, \TPAIR\ (\TOPERATIONLIST) \TYS) \STACKCONCAT\EMPTYSTACK\ be  the final stack, where \VOPERATIONLIST\ repesents a operation list

\begin{figure} []
\begin{align*}
T, U &::= \\
   &\Mid\ \langle \text{comparable type}\rangle \\
   &\Mid\ \text{option}\ \langle\text{type}\rangle \\
   &\Mid\ \text{list}\ \langle\text{type}\rangle \\
   &\Mid\ \text{set}\ \langle\text{comparable type}\rangle \\
   &\Mid\ \text{operation} \\
   &\Mid\ \text{contract}\ \langle\text{type}\rangle \\
   &\Mid\ \text{ticket}\ \langle\text{comparable type}\rangle \\
   &\Mid\ \text{pair}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{or}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{lambda}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{map}\ \langle\text{comparable type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{big-map}\ \langle\text{comparable type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{bls12-381-g1} \\
   &\Mid\ \text{bls12-381-g2} \\
   &\Mid\ \text{bls12-381-fr} \\
   &\Mid\ \text{sapling-transaction}\ \langle\text{natural number constant}\rangle \\
   &\Mid\ \text{sapling-state}\ \langle\text{natural number constant}\rangle \\
   &\Mid\ \text{chest} \\
   &\Mid\ \text{chest-key} \\
\langle\text{comparable type}\rangle ::= \\
   &\Mid\ \text{unit} \\
   &\Mid\ \text{never} \\
   &\Mid\ \text{bool} \\
   &\Mid\ \text{int}\\
   &\Mid\ \text{nat}\\
   &\Mid\ \text{string}\\
   &\Mid\ \text{chain-id}\\
   &\Mid\ \text{bytes}\\
   &\Mid\ \text{mutez}\\
   &\Mid\ \text{key-hash}\\
   &\Mid\ \text{key}\\
   &\Mid\ \text{signature}\\
   &\Mid\ \text{timestamp}\\
   &\Mid\ \text{address}\\
   &\Mid\ \text{tx-rollup-l2-address}\\
   &\Mid\ \text{option}\ \langle\text{comparable type}\rangle\\
   &\Mid\ \text{or}\ \langle\text{comparable type}\rangle \langle\text{comparable type}\rangle\\
   &\Mid\ \text{pair}\ \langle\text{comparable type}\rangle \langle\text{comparable type}\rangle \DOT \\
\end{align*}
\caption{Types}
\label{fig:type}
\end{figure}

\begin{figure}[h]
\begin{align*}
\text{t} &::= \\
   &\Mid\ \langle \text{variable} \rangle \\
   &\Mid\ \langle \text{account constant} \rangle \\
   &\Mid\ \langle \text{int constant} \rangle \\
   &\Mid\ \langle \text{string constant} \rangle \\
   &\Mid\ \langle \text{byte sequence constant} \rangle \\
   &\Mid\ \UNIT \\
   &\Mid\ \TRUE \\
   &\Mid\ \FALSE \\
   &\Mid\ \PAIR\ \text{t1 t2}\\
   &\Mid\ \LEFT\ \text{t}\\
   &\Mid\ \RIGHT\ \text{t}\\ 
   &\Mid\ \SOME\ \text{t}\\
   &\Mid\ \NONE \\
   &\Mid\ \text{\{t ; ... \}}\\
   &\Mid\ \text{\{ Elt t1 t2 ; ... \}}\\
   &\Mid\ \{ \langle\text{instruction}\rangle; ... \}   \\
\langle \text{variable} \rangle &::= \\ 
   &\Mid\ \VAR\\
\langle \text{account constant} \rangle &::= \\ 
   &\Mid\ \text{balance} \\
   &\Mid\ \text{amount} \\
   &\Mid\ \text{sender} \\
   &\Mid\ \text{source} \\
   &\Mid\ \text{now} \\
   &\Mid\ \text{level} \\
   &\Mid\ \text{chain-id} \\
   &\Mid\ \text{self}  \\
   &\Mid\ \text{self-address}  \\
   &\Mid\ \text{total-voting-power}  \\
   &\Mid\ \text{voting-power}  \\
\langle \text{natural number constant} \rangle &::= \\ 
   &\Mid\ \text {[0-9]+} \\
\langle \text{int constant} \rangle &::= \\
  &\Mid\ \langle \text{natural number constant} \rangle \\
  &\Mid\ \text{-} \langle \text{natural number constant} \rangle \\
\langle\text{string constant}\rangle &::= \\
  &\Mid\ \text{"} \langle \text{string content}\rangle\text{*"} \\
\langle\text{instruction}\rangle &::= \\
  &\Mid\ \DROP \\
  &\Mid\ \DROP \langle\text{natural number constant}\rangle \\
  &\text{...}
\end{align*}
\caption{Terms}
\label{fig:term}
\end{figure}

\begin{figure}[h]
\begin{align*}
\text{p} &::= \\
   &\Mid\ \langle \text{atomic formula} \rangle \\
   &\Mid\ \NEG\ \text{p} \
   \Mid\ \text{p}\ \CNF\ \text{q} \
   \Mid\ \text{p}\ \DNF\ \text{q} \\
\langle \text{atomic formula} \rangle &::= \\ 
    &\Mid\ \langle \text{butop} \rangle \langle \text{bterm} \rangle\\ 
   &\Mid\  \langle \text{bterm} \rangle  \langle \text{biop} \rangle \langle \text{bterm} \rangle\\ 
\langle \text{butop} \rangle &::= \\
   &\Mid\ \text{not} \\
\langle \text{biop} \rangle &::= \\
   &\Mid\ \text{$=$} \
   \Mid\ \text{$>$} \
   \Mid\ \text{$<$} \
   \Mid\ \text{$>=$} \
   \Mid\ \text{$=<$} \
   \Mid\ \text{$!=$} \\
   &\Mid\ \text{and}\ \Mid\ \text{or}\ \Mid\ \text{xor}\\
\langle \text{bterm} \rangle &::= \\ 
   &\Mid\ \text{t} \\
   &\Mid\ \langle \text{unop} \rangle \text{t} \\ 
   &\Mid\ \text{t}\ \langle \text{binop} \rangle \text{t}\\  
\langle \text{unop} \rangle &::= \\ 
   &\Mid\ \text{abs} \\
   &\Mid\ \text{size} \\
   &\Mid\ \text{int} \\
   &\Mid\ \text{contract}  \langle \text{type} \rangle \\
   &\Mid\ \text{isleft} \\
   &\Mid\ \text{isnone} \\
   &\Mid\ \text{neg} \\
   &\Mid\ \text{blake2b} \\
   &\Mid\ \text{hash-key} \\
   &\Mid\ \text{keccak} \\
   &\Mid\ \text{pairing-check} \\
   &\Mid\ \text{sha256} \\
   &\Mid\ \text{sha3} \\
   &\Mid\ \text{sha512} \\
   &\Mid\ \text{implicit-account} \\
   &\Mid\ \text{check-sig} \\
\langle \text{bitop} \rangle &::= \\ 
   &\Mid\ \text{+}\ \Mid\ \text{-}\ \Mid\ \text{*}\ \Mid\ \text{/}
\end{align*}
\caption{Predicates}
\label{fig:predicate}
\end{figure}

%\paragraph{Sequence actions.}
%\begin{figure}[h]
%\center
%\includegraphics[width= 10 cm]{life-cycle-1}
%\caption{Life-cycle for the auction contract.}
%\label{fca1}
%\end{figure}

\pagebreak
\subsection{Rules}
The rule semantic is defined by several kinds of transitions:
\begin{enumerate}
\item \ExprTrans\ single-step evaluation of an expression in a system state,
\item \StateTrans\ internal transitions of a system state, which is non-deterministic  as shown for instructions \ITER\ and \LOOP,
\item \SystemTrans\ symbolic system transitions.
\end{enumerate}

\subsubsection{System rules}
\begin{mathpar}
\inferrule[INVALID-PRE]
  { \NEG\ \PREDICATE
  }{
  \{[\INSTRUCTION, \STACK, \PREDICATE]\} \cup \SYSTEM \SystemTrans \SYSTEM}
\end{mathpar}

\subsubsection{Instruction rules}
\paragraph{Control structures}
%EXE
\begin{mathpar}
  \inferrule[EXEC]
  {
    [\INSTRUCTIONONE, (\StackOne, \TYF) \STACKCONCAT \EMPTYSTACK, 
    Q]
    \StateTrans^*
    [ \EMPTYSTACK, (\StackOne', \TYS) \STACKCONCAT Q']
  }{
     [(\EXEC; \INSTRUCTION),   (\{\INSTRUCTIONONE\}, \TYF\ \rightarrow\ \TYS) \STACKCONCAT
    (\StackOne, \TYF) \STACKCONCAT \STACK, \PREDICATE \wedge Q] 
    \StateTrans \
    [ \INSTRUCTION, (\StackOne', \TYS) \STACKCONCAT \STACK,
    \PREDICATE \wedge Q']}
\end{mathpar}

%APPLY
\begin{mathpar}
  \inferrule[APPLY]
  {
  }{
    [(\APPLY; \INSTRUCTION), (\StackOne, \TYF) \STACKCONCAT(\{\INSTRUCTIONONE\}, \TLAMBDA\ (\TPAIR\ \TYF\  \TYS)\ \TYT) \STACKCONCAT\STACK, \PREDICATE] \StateTrans\\ [\INSTRUCTION, (\{\PUSH\ \TYF\ \StackOne; \IPAIR; \INSTRUCTIONONE\}, \TLAMBDA\ \TYS\ \TYT ) \STACKCONCAT\STACK, \PREDICATE]}
\end{mathpar}

%LAMBLA
\begin{mathpar}
  \inferrule[LAMBDA]
  {  
  }{
    [(\LAMBDA\ \TYF\ \TYS\ \{ \INSTRUCTIONONE \} ; \INSTRUCTION),\STACK, \PREDICATE] \StateTrans\ [\INSTRUCTION, (\{\INSTRUCTIONONE\}, \TLAMBDA\ \TYF\ \rightarrow\ \TYS) \STACKCONCAT\STACK, \PREDICATE]}
\end{mathpar}

%IF
\begin{mathpar}
  \inferrule[IF-true]
  {  
  }{
    [(\IF\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TBOOL) \STACKCONCAT\STACK, \PREDICATE]
    \StateTrans\
    [\INSTRUCTIONONE, \STACK, \PREDICATE\ \Wedge\ \StackOne]
  }

  \inferrule[IF-false]
  {  
  }{
    [(\IF\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TBOOL) \STACKCONCAT\STACK, \PREDICATE]
    \StateTrans\
   [\INSTRUCTIONTWO, \STACK, \PREDICATE\ \Wedge\ \NEG\
   \StackOne]
 }
\end{mathpar}

%IF-LEFT-LEFT
\begin{mathpar}
  \inferrule[IF-LEFT-left]
  {  
  }{
    [(\IFLEFT\ \INSTRUCTIONONE\ \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TOR\ \TYF\ \TYS) \STACKCONCAT \STACK, \PREDICATE]
    \StateTrans \
    [\INSTRUCTIONONE, (\VariableX, \TYF) \STACKCONCAT\STACK,
    \PREDICATE \wedge (\StackOne\ \EQUAL\ \LEFT\ \VariableX)]
  }
\end{mathpar}

%IF-LEFT-RIGHT
\begin{mathpar}
  \inferrule[IF-LEFT-right]
  {  
  }{
    [(\IFLEFT\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),
    (\StackOne, \TOR\ \TYF\ \TYS) \STACKCONCAT \STACK, \PREDICATE]
    \StateTrans \
    [\INSTRUCTIONTWO, (\VariableX, \TYS) \STACKCONCAT\STACK, \PREDICATE \wedge (\StackOne\ \EQUAL\ \RIGHT\ \VariableX))]
  }
\end{mathpar}

%IF_CONS
\begin{mathpar}
  \inferrule[IF-CONS-empty]
  {  
  }{
    [(\IFCONS\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),  (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE] \StateTrans\  [\INSTRUCTIONTWO; \INSTRUCTION, \STACK, \PREDICATE\ \Wedge\ (\StackOne\ \EQUAL\ \EMPTYLIST)]}
\end{mathpar}

\begin{mathpar}
  \inferrule[IF-CONS-nonempty]
  {  
  }{
   [(\IFCONS\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),  (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE], \SYSTEM\ \StateTrans\  \\ [\INSTRUCTIONONE, (\HEAD, \TY) \STACKCONCAT(\{ \TAIL \}, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ ( \StackOne\ \EQUAL\ \{\HEAD; \STAIL \})]}
\end{mathpar}

%IF_NON
\begin{mathpar}
  \inferrule[IF-NONE-none]
  {  
  }{
    [(\IFNONE\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),  (\StackOne, \TOPTION\ \TY) \STACKCONCAT\STACK, \PREDICATE], \SYSTEM\ \StateTrans\   [\INSTRUCTIONONE; \INSTRUCTION, \STACK, \PREDICATE\ \Wedge\ (\StackOne\ \EQUAL\ \NONE)]}
\end{mathpar}

\begin{mathpar}
  \inferrule[IF-NONE-some]
  {  
  }{
    [(\IFNONE\ \INSTRUCTIONONE\  \INSTRUCTIONTWO; \INSTRUCTION),   (\StackOne, \TOPTION\ \TY) \STACKCONCAT\STACK, \PREDICATE], \SYSTEM\ \StateTrans\   [\INSTRUCTIONTWO,  (\VariableX, \TY) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ ( \StackOne\ \EQUAL\ \SOME\ \VariableX)]}
\end{mathpar}

%LOOP
\begin{mathpar}
  \inferrule[LOOP-true]
  {  
  }{
    [(\LOOP\ \INSTRUCTIONONE; \INSTRUCTION),  (\StackOne, \TBOOL)
    \STACKCONCAT\STACK, \PREDICATE]
    \StateTrans\
    [(\INSTRUCTIONONE; \LOOP\ \INSTRUCTIONONE; \INSTRUCTION),
    \STACK, \PREDICATE \wedge \StackOne]
  }

  \inferrule[LOOP-false]
  {  
  }{
    [(\LOOP\ \INSTRUCTIONONE; \INSTRUCTION),  (\StackOne, \TBOOL) \STACKCONCAT\
    \STACK, \PREDICATE]
    \StateTrans\
   [\INSTRUCTION, \STACK, \PREDICATE \wedge
   (\NEG\StackOne)]
   }
\end{mathpar}

%%ITER
\begin{mathpar}
  \inferrule[ITER-empty]
  {  
  }{
    [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE] \StateTrans\ [\INSTRUCTION, \STACK,  \PREDICATE\ \Wedge\  (\StackOne\ \EQUAL\ \EMPTYLIST)]
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[ITER-nonempty]
  {  [\ITER,  (\HEAD, \TY) \STACKCONCAT\STACK, 
    Q]
    \StateTrans^*
    [ \EMPTYSTACK,  \STACK', Q']
  }{
    [(\ITER\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE \wedge Q] \StateTrans \\ [(\TITER\ \INSTRUCTIONONE ; \INSTRUCTION), \{\STAIL\}\ \STACKCONCAT\STACK',  \PREDICATE\ \Wedge  ( \StackOne\ \EQUAL\ \{\HEAD; \STAIL \})  \wedge Q']
  }
\end{mathpar}

\paragraph{Stack Manipulation}
%DIG
\begin{mathpar}
\inferrule[DIG]
  {
   \FLEN(\A) \EQUAL\ \N
  }
  {[(\DIG\ \N ; \INSTRUCTION), \A\ \At\ (\StackOne, \TY) \STACKCONCAT\B, \PREDICATE] \StateTrans 
[\INSTRUCTION, (\StackOne, \TY) \STACKCONCAT\A\ \At\ \B, \PREDICATE]}
\end{mathpar}

%DIP
\begin{mathpar}
  \inferrule[DIP]
  {
    [\INSTRUCTIONONE,  \STACK, Q]
    \StateTrans^*
    [\EMPTYSTACK,  \STACK_1, Q']
  }
  {[(\DIP\ \INSTRUCTIONONE; \INSTRUCTION), (\StackOne, \TY) \STACKCONCAT
    \STACK, \PREDICATE \wedge Q]
    \StateTrans 
    [\INSTRUCTION, (\StackOne, \TY) \STACKCONCAT\STACK_1,
    \PREDICATE \wedge Q']
  }
\end{mathpar}

%DIP n
\begin{mathpar}
\inferrule[DIP n]
  { 
     \FLEN(\A) \EQUAL\ \N \\ [\INSTRUCTIONONE,  \B, Q]
    \StateTrans^*
    [\EMPTYSTACK,  \B_1, Q']
  }
  {[(\DIP\ \N\ \INSTRUCTIONONE; \INSTRUCTION), \A\ \At\ \B, \PREDICATE \wedge Q] \StateTrans 
[(\INSTRUCTION), \A\ \At\ \B_1, \PREDICATE \wedge Q']}
\end{mathpar}

%PUSH
\begin{mathpar}
  \inferrule[PUSH]
  {  
  }{
    [(\PUSH\ \TY\ \VariableX; \INSTRUCTION),\STACK, \PREDICATE] \StateTrans\ [\INSTRUCTION, (\VariableX, \TY) \STACKCONCAT\STACK, \PREDICATE]}
\end{mathpar}

\paragraph{Arithmetic operations}
%ADD
\begin{mathpar}
\inferrule[ADD]
  {
  }
  {[(\ADD; \INSTRUCTION), (\StackOne, \TNAT) \STACKCONCAT(\StackTwo, \TNAT) \STACKCONCAT\STACK, \PREDICATE] \StateTrans \
[\INSTRUCTION, (\VariableX, \TNAT) \STACKCONCAT\STACK, \PREDICATE \Wedge\ (\VariableX\ \EQUAL\ \StackOne\ \PLUS\ \StackTwo)]}
\end{mathpar}

%ABS
\begin{mathpar}
\inferrule[ABS]
  {
  }
  {
    [(\ABS; \INSTRUCTION), (\StackOne, \TINT) \STACKCONCAT \STACK,
    \PREDICATE]
    \StateTrans \
    [\INSTRUCTION, (\VariableX, \TNAT) \STACKCONCAT \STACK,
    \PREDICATE \wedge (\StackOne \ge 0 \Rightarrow \VariableX =
    \StackOne) \wedge (\StackOne <0 \Rightarrow \VariableX = -\StackOne)]
 }
\end{mathpar}

% COMPARE
\begin{mathpar}
\inferrule[COMPARE-nat]
  {
  }
  {
    [(\COMPARE ; \INSTRUCTION), (\StackOne, \TNAT) \STACKCONCAT (\StackTwo, \TNAT)
    \STACKCONCAT \STACK, \PREDICATE ]
    \SystemTrans \\
    [\INSTRUCTION, (\VariableX, \TINT) \STACKCONCAT \STACK, \PREDICATE
    \wedge\ (\StackOne > \StackTwo \Leftrightarrow \VariableX = 1)
    \wedge\ (\StackOne = \StackTwo \Leftrightarrow \VariableX = 0) 
    \wedge\ (\StackOne < \StackTwo \Leftrightarrow \VariableX = -1)]
    }
\end{mathpar}

\begin{mathpar}
\inferrule[COMPARE-some-some]
  {
  [\COMPARE,  (\VariableX, \TY) \STACKCONCAT (\VariableY, \TY) \STACKCONCAT\EMPTYSTACK, Q]
    \StateTrans^*
    [\EMPTYSTACK,  (\VariableA, \TINT) \STACKCONCAT\EMPTYSTACK, Q']
  }
  {
    [(\COMPARE ; \INSTRUCTION), (\StackOne, \TOPTION\ \TY) \STACKCONCAT (\StackTwo, \TOPTION\ \TY)
    \STACKCONCAT \STACK, \PREDICATE \wedge Q]
    \SystemTrans \\
    [\INSTRUCTION, (\VariableA, \TINT) \STACKCONCAT \STACK,  \PREDICATE
    \wedge (\StackOne\ \EQUAL\ \SOME\ \VariableX)
    \wedge (\StackTwo\ \EQUAL\ \SOME\ \VariableY)
    \wedge Q']
    }
\end{mathpar}

\begin{mathpar}
\inferrule[COMPARE-some-none]
  {
  }
  {
    [(\COMPARE ; \INSTRUCTION), (\StackOne, \TOPTION\ \TY) \STACKCONCAT (\StackTwo, \TOPTION\ \TY)
    \STACKCONCAT \STACK, \PREDICATE]
    \SystemTrans \\
    [\INSTRUCTION, (1, \TINT) \STACKCONCAT \STACK,  \PREDICATE
    \wedge (\StackOne\ \EQUAL\ \SOME\ \VariableX)
    \wedge (\StackTwo\ \EQUAL\ \NONE)
    ]
    }
\end{mathpar}

\paragraph{Boolean operations}
%XOR
\begin{mathpar}
\inferrule[XOR]
  {
  }
  {[(\XOR; \INSTRUCTION), (\StackOne, \TBOOL) \STACKCONCAT(\StackTwo, \TBOOL) \STACKCONCAT\STACK, \PREDICATE] \StateTrans \
[\INSTRUCTION, (\VariableX, \TBOOL) \STACKCONCAT\STACK, \PREDICATE \Wedge\ (\VariableX\ \EQUAL\ \StackOne\ \FXOR\ \StackTwo)]}
\end{mathpar}

\paragraph{Crytographic oprerations}
%HASH-KEY
\begin{mathpar}
\inferrule[HASH-KEY]
  {
  }
  {[(\HASHKEY; \INSTRUCTION), (\StackOne, \TBYTE) \STACKCONCAT\STACK, \PREDICATE] \StateTrans \
[\INSTRUCTION, (\VariableX, \TBYTE) \STACKCONCAT\STACK, \PREDICATE \Wedge\ (\VariableX\ = \FHASHKEY(\StackOne))]}
\end{mathpar}

%AMOUNT
\paragraph{Blockchain operations}
\begin{mathpar}
\inferrule[AMOUNT]
  {
  }
  {[(\AMOUNT; \INSTRUCTION), \STACK, \PREDICATE] \StateTrans \
[\INSTRUCTION, (\VAMOUNT, \TMUTEZ) \STACKCONCAT\STACK, \PREDICATE}
\end{mathpar}

%CONTRACT ty
\begin{mathpar}
\inferrule[CONTRACT TY - some]
  {
  }
  {[(\CONTRACT\ \TY ; \INSTRUCTION), (\StackOne, \TADDR) \STACKCONCAT\STACK, \PREDICATE] \SystemTrans \
[\INSTRUCTION, (\SOME\ \VariableX, \TOPTION\ (\TCONTRACT\ \TY)) \STACKCONCAT\STACK, \\ \PREDICATE \Wedge\ (\GETCONTRACTTYPE(\StackOne, \TY) = \SOME\ \VariableX)]}
\end{mathpar}

\begin{mathpar}
\inferrule[CONTRACT TY - none]
  {
  }
  {[(\CONTRACT\ \TY ; \INSTRUCTION), (\StackOne, \TADDR) \STACKCONCAT\STACK, \PREDICATE] \SystemTrans \
[\INSTRUCTION, \NONE \STACKCONCAT\STACK, \PREDICATE \Wedge\ (\GETCONTRACTTYPE(\StackOne, \TY) = \NONE]}
\end{mathpar}

\paragraph{Operations on data structures}
%CAR
\begin{mathpar}
\inferrule[\CAR]
  {
  }
  {[(\CAR; \INSTRUCTION), (\StackOne, \TPAIR\ \TYF\ \TYS) \STACKCONCAT\STACK, \PREDICATE] \StateTrans \
[\INSTRUCTION, (\VariableX, \TYF) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ (\StackOne\ \EQUAL\ \PAIR\ \VariableX\ \VariableY)]}
\end{mathpar}

\begin{mathpar}
\inferrule[CONCAT]
  {
  }
  {[(\CONCAT; \INSTRUCTION), (\StackOne, \TYLIST\ \TSTR) \STACKCONCAT\STACK,  \PREDICATE] \StateTrans 
[\INSTRUCTION,  ("",  \TSTR) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ (\StackOne\ \EQUAL\ \EMPTYLIST)]}
\end{mathpar}

\begin{mathpar}
\inferrule[CONCAT]
  {
  [\CONCAT,  (\{\TAIL\}, \TYLIST\ \TSTR) \STACKCONCAT\EMPTYSTACK, Q]
    \StateTrans^*
    [\EMPTYSTACK, (\StackTwo,  \TSTR) \STACKCONCAT\EMPTYSTACK, Q']
  }
  {[(\CONCAT; \INSTRUCTION),  (\StackOne, \TYLIST\ \TSTR) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ Q] \StateTrans \\
[\INSTRUCTION, ( \HEAD\ \STRINGCONCAT\ \StackTwo, \TSTR) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ (\StackOne\ \EQUAL\ \{\HEAD; \TAIL\}) \Wedge\ Q']}
\end{mathpar}

%MEM
\begin{mathpar}
\inferrule[MEM-empty]
  {
  }
  {[(\MEM; \INSTRUCTION), (\StackOne, \TYF) \STACKCONCAT(\StackTwo, \TYMAP\ \TYF\ \TYS) \STACKCONCAT\STACK, \PREDICATE] \StateTrans \
[\INSTRUCTION, (\FALSE, \TBOOL) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ (\StackTwo\ \EQUAL\ \EMPTYSET)]}
\end{mathpar}

\begin{mathpar}
\inferrule[MEM-nonempty]
  {
  [\COMPARE, (\StackOne, \TYF) \STACKCONCAT (\K, \TYF) \STACKCONCAT\EMPTYSTACK, Q]
    \StateTrans^*
    [\EMPTYSTACK,  (\VariableB, \TINT) \STACKCONCAT\EMPTYSTACK, Q']
  }
  {[(\MEM; \INSTRUCTION), (\StackOne, \TYF) \STACKCONCAT(\StackTwo, \TYMAP\ \TYF\ \TYS) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ Q] \StateTrans  \
[\MEM; \INSTRUCTION, \StackOne\ \STACKCONCAT\{\LESS\ \M\ \MORE\} \STACKCONCAT\STACK, \\ \PREDICATE\ \Wedge\ Q' \Wedge\ (\StackTwo\ \EQUAL\ \{\ELT\ \K\ \V; \LESS\ \M\ \MORE\})  \Wedge\ (\VariableB\ \EQUAL\ \ONE)]}
\end{mathpar}

\begin{mathpar}
\inferrule[MEM-nonempty]
  {
  [\COMPARE, (\StackOne, \TYF) \STACKCONCAT (\K, \TYF) \STACKCONCAT\EMPTYSTACK, Q]
    \StateTrans^*
    [\EMPTYSTACK,  (\VariableB, \TINT) \STACKCONCAT\EMPTYSTACK, Q']
  }
  {[(\MEM; \INSTRUCTION), (\StackOne, \TYF) \STACKCONCAT(\StackTwo, \TYMAP\ \TYF\ \TYS) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ Q] \StateTrans  \\
[\INSTRUCTION, (\TRUE, \TBOOL) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ Q' \Wedge\ (\StackTwo\ \EQUAL\ \{\ELT\ \K\ \V; \LESS\ \M\ \MORE\}) \Wedge\ (\VariableB\ \EQUAL\ \ZERO)]}
\end{mathpar}

\begin{mathpar}
\inferrule[MEM-nonempty]
  {
  [\COMPARE, (\StackOne, \TYF) \STACKCONCAT (\K, \TYF) \STACKCONCAT\EMPTYSTACK, Q]
    \StateTrans^*
    [\EMPTYSTACK,  (\VariableB, \TINT) \STACKCONCAT\EMPTYSTACK, Q']
  }
  {[(\MEM; \INSTRUCTION), (\StackOne, \TYF) \STACKCONCAT(\StackTwo, \TYMAP\ \TYF\ \TYS) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ Q] \StateTrans  \\
[\INSTRUCTION, (\FALSE, \TBOOL) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ Q' \Wedge\ (\StackTwo\ \EQUAL\ \{\ELT\ \K\ \V; \LESS\ \M\ \MORE\})  \Wedge\ (\VariableB\ \EQUAL\ \MINUS\ \ONE)]}
\end{mathpar}

%MAP
\begin{mathpar}
\inferrule[MAP-empty]
  {
  }
  {[(\MAP\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE] \StateTrans \
[\INSTRUCTION, (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ (\StackOne\ \EQUAL\ \EMPTYLIST)]}
\end{mathpar}

\begin{mathpar}
\inferrule[MAP-nonempty]
  {
  [\INSTRUCTIONONE, (\HEAD, \TY) \STACKCONCAT\EMPTYSTACK, Q1]
    \StateTrans^*
    [\EMPTYSTACK,  (\PHEAD, \TY) \STACKCONCAT\EMPTYSTACK , Q1'] \\  [\MAP\ \INSTRUCTIONONE, (\{\TAIL\}, \TYLIST\ \TY) \STACKCONCAT\EMPTYSTACK , Q2]
    \StateTrans^*
    [\EMPTYSTACK,  (\{\PTAIL\}, \TYLIST\ \TY) \STACKCONCAT\EMPTYSTACK , Q2']
  }
  {[(\MAP\ \INSTRUCTIONONE ; \INSTRUCTION), (\StackOne, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ Q1 \Wedge\ Q2 ] \StateTrans  \\
[\INSTRUCTION, (\{\PHEAD; \PTAIL\}, \TYLIST\ \TY) \STACKCONCAT\STACK, \PREDICATE\ \Wedge\ Q1' \Wedge\ Q2'  \Wedge\ (\StackOne \EQUAL\ \{\HEAD; \TAIL\})]}
\end{mathpar}
